---
title: "Do you even lift? In any case, know if you're doing it properly"
author: "EGo"
date: "06/06/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Loading relevant libraries
```{r libraries, results='hide'}
library(caret)
library(tidyverse)
library(ranger)
```

# Step 1: Get the data.
- We have the URLs. You know the drill:
```{r getData, cache=TRUE}
urlTrain <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"
urlTest <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"

if(!file.exists("pmlTrain.csv")){
        download.file(urlTrain,
                      destfile="pmlTrain.csv")
}

if(!file.exists("pmlTest.csv")){
        download.file(urlTest,
                      destfile="pmlTest.csv")
}

pmlTrain <- read.csv("pmlTrain.csv",
                     header=TRUE)
pmlTest <- read.csv("pmlTest.csv",
                    header=TRUE)
```

#Step 2: Clean the data
- There are many NA columns. Though not entirely NA, they are mostly NA with only about 2% non-NAs.
```{r removeNAs}
NAs <- which(is.na(pmlTrain[1, ]))
training <- pmlTrain[, -NAs]
```
- Near zero variance predictors are mostly the same for each outcome value; they don't add much. Let's exclude them too.
```{r removeNZVs, cache=TRUE}
nzv <- nearZeroVar(training)
training <- training[, -nzv]
```
- The first few columns aren't adding usable info. They're metadata. Off with them!
```{r trimVars}
training <- training %>% select(-(1:6))
```
- Some columns are of the integer class. Let's change them to numerics. It's easier to deal with that.
```{r makeNumeric}
print(sapply(training, class))
nonNumerics <- c(1, 5, 9:14, 18, 22:27, 31, 35:39, 44, 48:51)
for(i in nonNumerics){
        training[, i] <- as.numeric(training[, i])
}
```

# Step 3: Fit the model
- We have a lot of variables still. Let's whittle them down using PCA—capturing just 80% variance; we don't want to overfit.
```{r PCA}
preProc <- preProcess(training[, -53], method="pca", thresh=0.8)
trainPC <- predict(preProc, training[, -53])
```
- A classification prediction problem? Random forest sounds good. Using the {ranger} method here.
```{r fitModel, cache=TRUE}
set.seed(2020)
ptm2 <- proc.time()
fit <- train(x=trainPC, y=training$classe, method="ranger",
              trControl=trainControl(method="repeatedcv", verboseIter=F))
confusionMatrix(predict(fit, trainPC), training$classe)$overall
print(proc.time()-ptm2)
```
- We have cross-validated. Let's look at the OOB error. It's quite low.
```{r OOBerror}
fit$finalModel
```

# Step 4: Predict on the test set
- We need to preprocess the testing set first
```{r testProcess}
testing <- pmlTest[, -NAs]
testing <- testing[, -nzv]
testing <- testing %>% select(-(1:6))
nonNumerics <- c(1, 5, 9:14, 18, 22:27, 31, 35:39, 44, 48:51)
for(i in nonNumerics){
        testing[, i] <- as.numeric(testing[, i])
}
testPC <- predict(preProc, testing[, -53])
```
- And finally, we can predict!
```{r predict}
print(predict(fit, testPC))
```